---
title: " Tidy PMS 93 00 07 Data"
author: "M K Jamieson"
date: 
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
# load libraries 
pacman::p_load(knitr, 
               tidyverse, 
               haven,
               dplyr,
               dataMaid,
               janitor,
               dlookr)
```

# Cleaning Data

## Read in Source Data
```{r read in data}
# read in 1993
source_pms93 <- haven::read_spss("pms93.por", 
                                user_na = TRUE) %>%
  as_factor()



# read in 2000
source_pms00 <- haven::read_sav("pms00.sav", 
                                user_na = TRUE) %>%
  as_factor()



# read in 2007
source_pms07 <- haven::read_sav("pms07.sav", 
                                user_na = TRUE) %>%
  as_factor() 
```

## Subset & Check 2007 Data
```{r subset 07}
#### variables of interest subset 2007 ####
subset_pms07 <- source_pms07 %>%
  dplyr::select("pserial","ResSex","Health6","Language","Age10yr",
                "Age20yr","ETHNIC4","Origin","EDQUAL5","AnyQuals",
                "HiQuals","ResMarDF",
                "DVILO3a","DVPrac","Care1","Care2","Care3","Care4",
                "Care5","Everwk","yearjbl","JbReas","Looked",
                "YInAct","LookStop","LookSto2","Stat","Solo",
                "HrsWeek","PTWkHour","EmpStY","LookNot1","LookNot2",
                "LookNot3","DVLastWk","NotWk","WkShel1","HrsWork",
                "LookSto3","DiffJob","WkShel2","WkShel3","SEmpStY",
                "LookNow","LookAtAl","JobstM","EmpNo","Manage",
                "FtPtWk","NotWk","SEG","SC","SrcInc1","SrcInc2",
                "SrcInc3","SrcInc4","SrcInc5","SrcInc6","Gross4",
                "Gross4a","Ten1","LLord","SF6","SF7","SF9","SF11",
                "SF12","CONHOMD","IMon","FollUpGrp",
                "eligible","Stage2","Origin","DVPrac","Care1","Care2",
                "Care3","Care4","Care5","Medic","CnslHav","CnslTak",
                "CnslLng","DocYear","DocPsyc","PMatNum","DocTalk",
                "DocWeeks","InWhy","OutWhy","DayY","DayWht1",
                "DayWht2","DayWht3","DayWht4","CC2aY",
                "CC2Y1","CC2Y2","CC2Y3","CC2Y5","CC2Y6","CC2Y7",
                "CC2Y8","CC2Y9","MentHos","Ten1","Gross4a","qimd",
                "gor06","newsha","SF1","Happy","tenure","newten",
                "hhdtype","DISsex","DISeth","DISrel","DISAge",
                "DISmen","DISphy",
                "DISsori","CISRFOUR","panic","GAD","madd","OCD",
                "phob","dep","neurotic","nosymp","numdis",
                "PTSDever","MajorT16","PTSDcom","PTSDPos",
                "suicthwk","suicthyr","suicthlf","DSHlife",
                "suicatwk","suicatyr","suicatlf","DSHtry","suihfri",
                "suihhos","suihcom","suihhel","suihoth","DSH5",
                "DSHharm","DSH9","DSH10","PsycProb","Psycdis",
                "Psycdis_wt","BPDPH2","bdpd_wt","ASPDPH2","aspd_wt",
                "DVADHD1","DVADHD2","DVADHD4","scoff2","EDImpact",
                "bmigp4","DVAudit1","AUDITgp","DRNKPROB","SADQCSC",
                "SADQGP","AUDSAD2","Cannever","Cannyear","AMPHEVER",
                "AMPHYEAR","COCAEVER","CRACEVER","CRACYEAR",
                "ECSTEVER","ECSTYEAR","HEROEVER","HEROYEAR",
                "ACIDEVER","ACIDYEAR","MUSHEVER","MUSHYEAR",
                "METHEVER","METHYEAR","TRANEVER","TRANYEAR",
                "AMYLEVER","AMYLYEAR","ANABEVER",
                "ANABYEAR","GLUEEVER","GLUEYEAR","DRUGEVER",
                "DRUGYEAR","DRUGDEP",
                "DRUGDEP2","GamYR","dsmscgr","dsmprob","dsmpath",
                "numdiag","numdiag4","LCA","cluster1","cluster2",
                "cluster3","cluster4","cluster5","cluster6",
                "MedPsyc","MedDep","MedHyp","MedAnx","MedADHD",
                "anymed","Doc2Wks","DocPsyc","INQTRMEN","OUTQTRME",
                "PSYCTHER","COGTHER","ARTTHER","SOCTRAIN","MARITHER"
                ,"SEXTHER","COUNSEL","OTHTHER","OTHTHER","anyther",
                "trtment","DAYCOMYR","PSYTRTYR","PSYLGTYR",
                "CPNYR","CLDNYR","OTHNSEYR","SOCWRKYR","CPNYR",
                "ERI3","SFQA","SFQB","SFQC","SFQD",
                "SFQE","SFQF","SFQG","SFQH","DLSS4","DLSS5",
                "PrimGrp","ChldInst","LACare","BothMaPa",
                "YNotBoth","MaOrPa","AnyChild","NoChild",
                "InDebt1","InDebt6",
                "SFHELPYR","HMHELPYR","OREACHYR","anyccar","anyhlca"
                ,"anydaca","wt_ints1","ipsu_0714","istrata_0714") %>%
  janitor::clean_names() %>%
  janitor::remove_empty("cols") %>%
  janitor::remove_empty("rows") 

head(subset_pms07, 5)

```

## Make Names Tidy for Variable Comparison
```{r arrange and lowercase for comparison}
# arrange alphebetically and make colnames lowercase

subset_pms07 <-subset_pms07[,order(colnames(subset_pms07))] %>% rename_all(tolower)
source_pms00 <-source_pms00[,order(colnames(source_pms00))] %>% rename_all(tolower)
source_pms93 <-source_pms93[,order(colnames(source_pms93))] %>% rename_all(tolower)

head(subset_pms07, 5)
head(source_pms00, 5)
head(source_pms93, 5)
```

## Looking for Common Variables in 2000 (with 2007)
+ manual search of data dictionary.
```{r common vars in 00 and 07}
#### Finding Common Variables Across Datasets ####
#common vars in 00 and 07
intersect(names(subset_pms07), names(source_pms00))

#colnames in 07 but not 00
setdiff(names(subset_pms07), names(source_pms00))
```

## Subset and Check 2000 Data
+ checking data dictionary 
```{r subset 00}
## subsetting for matching vars of interest 2000
subset_pms00 <- source_pms00 %>%
  dplyr::select("acidever","acidyear","age10yr","age20yr","amphever","amphyear",
                "amylever","amylyear","anabever","anabyear","anyquals",
                "artther","auditgp","audsad2","cannever","cannyear","cc2ay",
                "cc2y1","cc2y2","cc2y3","cc2y5","cc2y6","cc2y7","cc2y8","cc2y9",
                "chldinst","cisrfour","cldnyr","cnslhav","cnsllng","cnsltak",
                "cocaever","cogther","counsel","cpnyr", "cracever","cracyear",
                "daywht1","daywht2","daywht3","daywht4","dayy","dep","diffjob",
                "dlss4","dlss5","docpsyc","doctalk","docyear","drnkprob",
                "drugdep","drugdep2","drugever","drugyear","dsh10","dsh5","dsh9",
                "dvilo3a","dvlastwk","ecstever","ecstyear","edqual5","empno",
                "empsty","ethnic4","everwk","ftptwk","gad","glueever","glueyear",
                "gross4","gross4a","heroever","heroyear","hiquals","hmhelpyr",
                "hrsweek","hrswork","inqtrmen","inwhy","jbreas","jobstm","lacare","language","llord",   
                "lookatal","looked","looknot1","looknot2","looknot3","looknow",
                "looksto2","looksto3","manage","marither","menthos","methever",
                "methyear","mushever","mushyear","neurotic","nosymp","notwk",
                "ocd","oreachyr","origin","othnseyr","othther","outqtrme","outwhy",  
                "panic","phob","pmatnum","psycther","psylgtyr","psytrtyr","sadqgp","sc","seg","sempsty",
                "sexther","sf1","sf11","sf12","sf6","sf7","sf9","sfhelpyr",
                "soctrain","socwrkyr","solo","srcinc1","srcinc2","srcinc3",
                "srcinc4","srcinc5","srcinc6","stat","suicatlf","suicatwk",
                "suicatyr","suicthlf","suicthwk","suicthyr","ten1","tenure",
                "tranever","tranyear","trtment","wkshel1","wkshel2","yinact","respsex","sstrtreg","resmarst","dtjbl","jbreas","commany",
                "numchild","docwhat","treat","wt2","parents","mencenyr","othdayyr",
                "daycenyr","doctalk","suicthlf","suicatlf","q20","q23","b3a",
                "govreggb","numadult","pmatoday","indebt01","futrjbw1","mad",
                "psycdrug","drglong","vignass","schizo","psycho","drinknow") %>%
  janitor::clean_names() %>%
  janitor::remove_empty("cols") %>%
  janitor::remove_empty("rows") 
  

head(subset_pms00, 5)

```

## Looking for Common Variables in 1993 Data (with 2007)

```{r common vars in 93 and 00}
#common vars in 07 and 93
intersect(names(subset_pms07), names(source_pms93))
# in 00 and 93
intersect(names(subset_pms00), names(source_pms93))

#colnames in 07 but not 93
setdiff(names(subset_pms07), names(source_pms93))
#colnames in 00 but not 93
setdiff(names(subset_pms00), names(source_pms93))

```

## Subset and Check 1993 Data
+ data dictionary 
```{r subset 93}
## subset 1993 vars of interest
subset_pms93 <- source_pms93 %>%
  dplyr::select("age10yr","dep","ethnic4","gad","ocd","panic","phob",
                "sc","seg","tenure", "mad","eacid","espeed","eglue","be10d",
                "nchldren","a10","bb1a01","bb1d01","bb1eyr01","bb2",
                "bb2a01","ehash","bc2","bc3","bc4a01","bc4b01","bc4c01",
                "bc4d01m1","bb2bmt01","bb2byr01","bb2bne01","bb2cyr01",
                "bb2cmt01","bb2d01","ecoke","eheroin","eopium","eecstasy",
                "esleep","etranx","ab2","ab3a","bf2","bf4","bf8b","bf13",
                "region","area","hhold","sex01","age01","race01","a4","a5",
                "a5a","a5b","a11","a15","bg1","bg1a","bg1a1","bg1b","bg1c",
                "bg1d","bg1f","bg2","bg2d","bg2e","bg2f","bg2g","bg2h","bg2i",
                "bg2j","bg2k","bg3a","bg3b","bg4","bg4a","bf12a","bf12b","bf14","bf15a","bf16","bf17","bf17am1","bf18dna",
                "bf18","bf19","bf21","bf21a","bf21a1m1","bf22m1","bf24",
                "bf25","bb1a01","diagno11","subethn","submarst","workstat",
                "timeunem") %>%
  janitor::clean_names() %>%
  janitor::remove_empty("cols") %>%
  janitor::remove_empty("rows")
  

head(subset_pms93, 5)

```

# Exploring 

Plan
Univariate, tests of normality, visualising normality, frquencies, crosstabs, assosiations 
Bivariate, correlation, visulisation,

## Calculating Descriptives 93


```{r desc 93}



```
## Calculating Descriptives 00

```{r}

```


## Calculating Descriptives 07

```{r}

```


# Linear Regression 

```{r}

```

